### 2015-02-24 Studies/2012_PG_County/Report/PG_Co_Fire.tex
PG Co: fix minor typos from ERB

\chapter{Introduction}
\pagenumbering{arabic}
\setcounter{page}{1}
Part of the function of the Fire Research Division at NIST is to develop and apply technology, measurements, and standards to improve the understanding of the behavior, prevention, and control of fires to enhance fire fighting operations and equipment, fire suppression, fire investigations, and disaster response. NIST has previously used the Fire Dynamics Simulator (FDS) fire model to provide insight into the fire development and thermal conditions of fires that have resulted in line of duty deaths (LODDs)~\cite{Madrzykowski:1,Iowa,Texas,Bryner:Charleston,barowy:texas,Weinschenk:Chicago,Overholt:San_Francisco}. The objective of these studies has been to improve the safety and effectiveness of firefighters.

@@

As discussed in Section~\ref{geom}, the structure has dimensions of 9.14~m (30~ft) by 7.62~m (25~ft) and a total height of 7.31~m (24~ft). To sufficiently resolve the wind boundary condition, the computational domain was extended beyond the volume of the structure. The large domain led to the use of two separate grid resolutions, a 10~cm (3.9~in) resolution and a 20~cm (7.8~in) resolution. The 10~cm resolution encompassed the structure with a total volume 14~m (46~ft) by 12~m (39ft) by 8~m (26~ft) that required a total of 1.3 million computational cells. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10~cm. The 20~cm grid resolution extended the computational domain 5~m to the left and right sides and 4~m behind the 10~cm mesh to provide sufficient computational domain for the wind profile to develop around the structure. The mesh is coarser to cut down on computational time and because high accuracy in the fluid mechanics of the air flow far from the structure is not important in the analysis of this simulation. The 20~cm resolution resulted in an additional 216,000 computational cells.

Figure~\ref{fig:geom_grid} shows the front and rear of the structure and surrounding domain rendered in Smokeview with the 10~cm and 20~cm meshes. Note the finer mesh (smaller cells) surrounding the structure and the coarser mesh on each side. The domain was divided into 12 equally sized 10~cm meshes that surrounded the structure (each containing 150,000 grid cells) plus 4 equally sized 20~cm meshes to the left and right of the structure (each containing 32,000 grid cells) and one 20~cm mesh in the rear of the structure (containing 56,000 grid cells) for a total of 17 meshes. Multiple meshes allowed the simulation to be processed in parallel, which reduced the amount of required calculation time to approximately 3~days from over 1~week with a uniform 10~cm mesh. Figure~\ref{fig:mult_mesh} shows the entire structure within the computational domain that has been divided into multiple meshes. The boundary between the twelve 10~cm meshes and five 20~cm meshes is indicated by the {\bf bold} lines.

@@

A comparison of the differences between the HRR inside the structure (Fig.~\ref{fig:PG_Total_HRR}) and the total HRR (Fig.~\ref{fig:PG_9MW_HRR}) over the first 100 seconds of the simulation indicated that on average there was the equivalent of an approximate 0.7 MW fire(s) occurring through open external vents (basement windows). This aligns with observed post-fire damage and post-incident report~\cite{PGCounty2013}. Fig.~\ref{fig:ext_fire_dam} shows a post-fire image of the back left corner of the structure as well as a Smokeview rendering of the fire from the FDS simulation prior to the front door being opened by firefighters.

@@

After 100 seconds of simulation time, 5~\sfrac{1}{2}~minutes after the first 911 call, firefighters open the front door to make entry to the first floor of the structure. The change in ventilation to the structure impacted the HRR. After the door was opened combustion no longer took place outside of the structure, as shown in Fig.~\ref{fig:PG_Basement_9MW_HRR}, however the basement HRR was lower than the total HRR within the domain.

@@

The growth of the fire combined with a 8.9~m/s (20~mph) wind flowing normal to the rear of the structure caused the pressure to rise within the structure. Hot gases which may have built up within the structure will flow from high pressure to low pressure. Based on the event timeline, the front door was opened 100 seconds into the simulation, which would create a low pressure relief on the first floor. Due to firefighters entering the structure through the front door, the simulated pressure on the first floor prior to and just after opening the door is important. Figure~\ref{fig:pressure_slices} shows the calculated pressure conditions in the structure just before and after the front door was opened (100~s) in the simulation.

@@

At the estimated time of firefighter entry to the first floor (142~s into the simulation at 21:14:42; see Table~\ref{tab:combined_timeline}), the simulated flow magnitudes through the first floor (stairwell and front doors) remained similar to the conditions just after opening the front door. Velocities through both the doorway at the top of the stairs and the front door were approximately 13~m/s (29~mph), which was faster than the imposed wind speed (9~m/s or 20~mph) because of the addition of the buoyancy of the fire plume. The model temperatures increased through the stairwell doorway to between 400~$^{\circ}$C and 425~$^{\circ}$C (750-800~$^{\circ}$F) and through the front door to approximately 375~$^{\circ}$C (700~$^{\circ}$F). From the approximate timeline (Table~\ref{tab:combined_timeline}), the front door was forced closed due to airflow pushing against it, 207~s into the simulation. Figures~\ref{fig:flow_path_206s} and \ref{fig:flow_path_207s} show the conditions 1~s prior to and after the front door closed. The figures show that while the closed door stopped significant flow through the stairwell door and onto the first floor, the temperature hazard for the firefighters (temperatures at or above 400~$^{\circ}$C or 750~$^{\circ}$F) remained.

>>>>

\chapter{Introduction}
\pagenumbering{arabic}
\setcounter{page}{1}
Part of the function of the Fire Research Division at the National Institute of Standards and Technology (NIST) is to develop and apply technology, measurements, and standards to improve the understanding of the behavior, prevention, and control of fires to enhance fire fighting operations and equipment, fire suppression, fire investigations, and disaster response. NIST has previously used the Fire Dynamics Simulator (FDS) fire model to provide insight into the fire development and thermal conditions of fires that have resulted in line of duty deaths (LODDs)~\cite{Madrzykowski:1,Iowa,Texas,Bryner:Charleston,barowy:texas,Weinschenk:Chicago,Overholt:San_Francisco}. The objective of these studies has been to improve the safety and effectiveness of firefighters.

@@

As discussed in Section~\ref{geom}, the structure has dimensions of 9.14~m (30~ft) by 7.62~m (25~ft) and a total height of 7.31~m (24~ft). To sufficiently resolve the wind boundary condition, the computational domain was extended beyond the volume of the structure. The large domain led to the use of two separate grid resolutions, a 10~cm (3.9~in) resolution and a 20~cm (7.8~in) resolution. The 10~cm resolution encompassed the structure with a total volume 14~m (46~ft) by 12~m (39ft) by 8~m (26~ft) that required a total of 1.3 million computational cells. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10~cm. The 20~cm grid resolution extended the computational domain 5~m to the left and right sides and 4~m behind the 10~cm mesh to provide sufficient computational domain for the wind profile to develop around the structure. The mesh is coarser to cut down on computational time and because high accuracy in the fluid mechanics of the air flow far from the structure is not important in the analysis of this simulation. The 20~cm resolution resulted in an additional 216\,000 computational cells.

Figure~\ref{fig:geom_grid} shows the front and rear of the structure and surrounding domain rendered in Smokeview with the 10~cm and 20~cm meshes. Note the finer mesh (smaller cells) surrounding the structure and the coarser mesh on each side. The domain was divided into 12 equally sized 10~cm meshes that surrounded the structure (each containing 150\,000 grid cells) plus 4 equally sized 20~cm meshes to the left and right of the structure (each containing 32\,000 grid cells) and one 20~cm mesh in the rear of the structure (containing 56\,000 grid cells) for a total of 17 meshes. Multiple meshes allowed the simulation to be processed in parallel, which reduced the amount of required calculation time to approximately 3~days from over 1~week with a uniform 10~cm mesh. Figure~\ref{fig:mult_mesh} shows the entire structure within the computational domain that has been divided into multiple meshes. The boundary between the twelve 10~cm meshes and five 20~cm meshes is indicated by the {\bf bold} lines.

@@

A comparison of the differences between the HRR inside the structure (Fig.~\ref{fig:PG_Total_HRR}) and the total HRR (Fig.~\ref{fig:PG_9MW_HRR}) over the first 100~s of the simulation indicated that on average there was the equivalent of an approximate 0.7 MW fire(s) occurring through open external vents (basement windows). This aligns with observed post-fire damage and post-incident report~\cite{PGCounty2013}. Fig.~\ref{fig:ext_fire_dam} shows a post-fire image of the back left corner of the structure as well as a Smokeview rendering of the fire from the FDS simulation prior to the front door being opened by firefighters.

@@

After 100~s of simulation time, 5.5~min after the first 911 call, firefighters open the front door to make entry to the first floor of the structure. The change in ventilation to the structure impacted the HRR. After the door was opened combustion no longer took place outside of the structure, as shown in Fig.~\ref{fig:PG_Basement_9MW_HRR}, however the basement HRR was lower than the total HRR within the domain.

@@

The growth of the fire combined with a 8.9~m/s (20~mph) wind flowing normal to the rear of the structure caused the pressure to rise within the structure. Hot gases which may have built up within the structure will flow from high pressure to low pressure. Based on the event timeline, the front door was opened 100~s into the simulation, which would create a low pressure relief on the first floor. Due to firefighters entering the structure through the front door, the simulated pressure on the first floor prior to and just after opening the door is important. Figure~\ref{fig:pressure_slices} shows the calculated pressure conditions in the structure just before and after the front door was opened (100~s) in the simulation.

@@

At the estimated time of firefighter entry to the first floor (142~s into the simulation at 21:14:42; see Table~\ref{tab:combined_timeline}), the simulated flow magnitudes through the first floor (stairwell and front doors) remained similar to the conditions just after opening the front door. Velocities through both the doorway at the top of the stairs and the front door were approximately 13~m/s (29~mph), which was faster than the imposed wind speed (9~m/s or 20~mph) because of the addition of the buoyancy of the fire plume. The model temperatures increased through the stairwell doorway to between 400~$^{\circ}$C and 425~$^{\circ}$C (750~$^{\circ}$F to 800~$^{\circ}$F) and through the front door to approximately 375~$^{\circ}$C (700~$^{\circ}$F). From the approximate timeline (Table~\ref{tab:combined_timeline}), the front door was forced closed due to airflow pushing against it, 207~s into the simulation. Figures~\ref{fig:flow_path_206s} and \ref{fig:flow_path_207s} show the conditions 1~s prior to and after the front door closed. The figures show that while the closed door stopped significant flow through the stairwell door and onto the first floor, the temperature hazard for the firefighters (temperatures at or above 400~$^{\circ}$C or 750~$^{\circ}$F) remained.
